Indexed Containers

### 2011-05-10 journal/paper.tex
NeNf SPF

The initial algebra semantics of inductive types can be extended to
model inductive families by replacing functors on the category \ensuremath{\Conid{Set}}
with functors on the category of families indexed by a given type - in
the case of $\lambda$-terms this indexing type was \ensuremath{\Conid{Nat}}. The objects
of the category of families indexed over a type \ensuremath{\Conid{I}\;\in\;\Varid{set}} are
\ensuremath{\Conid{I}}-indexed families, i.e. functions of type \ensuremath{\Conid{I}\;\rightarrow\;\Conid{Set}}, and a
morphism between \ensuremath{\Conid{I}}-indexed families \ensuremath{\Conid{A},\Conid{B}\;\in\;\Conid{I}\;\rightarrow\;\Conid{Set}} is given by a
family of maps \ensuremath{\Varid{f}\;\in\;(\Varid{i}\;\in\;\Conid{I})\;\to \;\Conid{A}\;\Varid{i}\;\to \;\Conid{B}\;\Varid{i}}
Indeed, this category
is easily seen to be isomorphic to the slice category $\ensuremath{\Conid{Set}}/ \ensuremath{\Conid{I}}$ but
the chosen representation is more convenient type-theoretically. Using
$\Sigma$-types and equality types from Type Theory, we can define the
following endofunctors \ensuremath{\Conid{F}_{\Conid{Fin}},\Conid{F}_{\Conid{Vec}}} and \ensuremath{\Conid{F}_{\Conid{Lam}}}
on the category of families
over \ensuremath{\Conid{Nat}} whose initial algebras are \ensuremath{\Conid{Fin}} and \ensuremath{\Conid{Lam}}, respectively:

@@

>>>>

The initial algebra semantics of inductive types can be extended to
model inductive families by replacing functors on the category \ensuremath{\Conid{Set}}
with functors on the category of families indexed by a given type - in
the case of $\lambda$-terms this indexing type was \ensuremath{\Conid{Nat}}. The objects
of the category of families indexed over a type \ensuremath{\Conid{I}\;\in\;\Varid{set}} are
\ensuremath{\Conid{I}}-indexed families, i.e. functions of type \ensuremath{\Conid{I}\;\rightarrow\;\Conid{Set}}, and a
morphism between \ensuremath{\Conid{I}}-indexed families \ensuremath{\Conid{A},\Conid{B}\;\in\;\Conid{I}\;\rightarrow\;\Conid{Set}} is given by a
family of maps \ensuremath{\Varid{f}\;\in\;(\Varid{i}\;\in\;\Conid{I})\;\to \;\Conid{A}\;\Varid{i}\;\to \;\Conid{B}\;\Varid{i}}
Indeed, this category
is easily seen to be isomorphic to the slice category $\ensuremath{\Conid{Set}}/ \ensuremath{\Conid{I}}$ but
the chosen representation is more convenient type-theoretically. Using
$\Sigma$-types and equality types from Type Theory, we can define the
following endofunctors \ensuremath{\Conid{F}_{\Conid{Fin}}}, \ensuremath{\Conid{F}_{\Conid{Vec}}} and \ensuremath{\Conid{F}_{\Conid{Lam}}}
on the category of families
over \ensuremath{\Conid{Nat}} whose initial algebras are \ensuremath{\Conid{Fin}} and \ensuremath{\Conid{Lam}}, respectively:

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{FNe}\;\in\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{FNe}\;\Conid{X}\;\Conid{Y}\;\Varid{n}\;\mathrel{=}\;\Conid{Fin}\;\Varid{n}\;\uplus\;(\Conid{X}\;\Varid{n}\;\uplus\;\Conid{Y}\;\Varid{n}){}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Conid{FNf}\;\in\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{FNf}\;\Conid{X}\;\Conid{Y}\;\Varid{n}\;\mathrel{=}\;(\Conid{Y}\;\ensuremath{\mbox{$\circ$}}\;\Varid{suc})\;\Varid{n}\;\uplus\;\Conid{X}\;\Varid{n}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
